Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries

### 2018-12-31 docs/SupplementaryMaterial.Rmd
Made last changes and re-submitted

---
title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template_sup.tex
---

>>>>

---
title: "Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template_sup.tex
---

### 2018-12-31 docs/manuscript.Rmd
Made last changes and re-submitted

---
title: "Effectiveness of community-based TURF-reserves in small-scale fisheries"
rtitle: "Community-based TURF-reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"
institution: |
  $^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA \newline
  $^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA \newline
  $^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico \newline
  $^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "juancarlos@ucsb.edu"
abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial user rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last 12 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a 2014 regulation; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive support from the fishing communities. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, some of these reserves might be too young for the effects to show. Second, the reserves are not large enough to protect mobile species, like lobster. Third, variable and extreme oceanographic conditions have impacted harvested populations. Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches. However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

@@

TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean. They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks. These instruments can be combined by establishing a marine reserve within a TURF, thus making them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.

Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The benefits resulting from reserves established within TURFs (*i.e.* TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}. Although in theory these systems are successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success.

TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}. This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009}. There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}. However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.

Recent norms in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries. In Mexico, a norm created in 2014 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}). Since 2012, old and new marine reserves have gained legal recognition as Fishing Refuges. Of these, 18 were originally implemented as within TURFs. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico. These three case studies span a range of ecological and social conditions representative of different regions of Mexico. The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness. We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild their coastal fisheries.

@@

In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue (*i.e.* Fish Refuges; \citet{nom}). Fish Refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years). After these five years have passed, fishers have the opportunity to open the reserves to fishing or continue to have them. Our work focuses on Fish Refuges implemented as community-based TURF-reserves that occur in small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A). Each TURF is elusively fished by one cooperative, and each community usually hosts no more than one cooperative. The profits from each TURF are shared amongst all fishers from the cooperative. Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (*i.e.* TURF-reserves). Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges following a series of studies outlined in the regulation \citep{nom}.

@@

We evaluate three community-based no-take TURF-reserves implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A). The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds to protect commercially important invertebrate species; mainly lobster and abalone. These reserves obtained legal recognition in 2018 \citep{dof_website_2018}.

The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the *Cozumel* fishing cooperative; Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a local community. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and re-enactments \citep{dof_website_2017b}.

@@

The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have biological or socioeconomic benefits \citep{nom}. For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation. However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place. We therefore incorporate a reduced version of the Social Ecological Systems framework \citep{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries. The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop. The following two sections describe our data collection methods and analyses of biological and socioeconomic indicators as well as the SES analysis.

@@

We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control sites. Reserve sites are those within the reserves, and thus no fishing takes place. Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) Are not directly adjacent to the reserves. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \citep{francinifilho_2008,moland_2013,Villasenor-Derbez_2018}.

@@

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves. In this case our treated unit are the cooperatives (*i.e.* communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve. Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}. A table with summary statistics for this data is provided in the supplementary materials **(Table S3, Figure S5)**.

Data for the evaluation of the SES were collected at the community-level from official documents used in the creation and designation of the marine reserves. These includes the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also complimented information based on the authors' experience and knowledge of the communities. We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}). The next sections further describe the variables and indicators chosen for each of these.

@@

\begin{equation}
I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t,j}
\label{eqn:reg_bio}
\end{equation}

Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively). Therefore, we would expect this term to be positive if the indicator increases because of the reserve implementation. Finally, $\epsilon_{i,t,j}$ represents the error term of the regression.

Socioeconomic indicators are evaluated with a similar approach. Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants. Each focal community (Isla Natividad and Maria Elena) has three counterfactual communities.

\begin{equation}
I_{i,t,j} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t,j}
\label{eqn:soc_reg}
\end{equation}

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$). These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}. All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All analyses were performed in R `r paste0("version ", version$major, ".", version$minor)` and R Studio version 1.1.456 \citep{R_2018}. All data and code needed to reproduce our analyses are available in a [GitHub repository](https://github.com/jcvdav/ReserveEffect).

@@

While many ecology studies have used BACI sampling designs and respective analyses (*e.g.* \citet{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (*e.g.* \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness. For example, we capture previously described patterns like the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig. \ref{fig:indicators}A), a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish. Yet, our empirical approach assumes control sites are a plausible counterfactual for treated sites. This implies that treated sites would have followed the same trend as control sites, had the reserves not been implemented. Nonetheless, temporal trends for each site don't show any significant increases (Figures S1-S4), supporting our findings of lack of change in the indicators used.

Our analyses of socioeconomic indicators has two limitations. First, we only look at landings and revenues by landings for communities with and without TURF-reserves. There are a number of other possible indicators that could show a change due to the implementation of the reserve. Notably, one often cited in the literature is additional benefits, such as tourism \citep{viana_2017}. However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism. A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve. For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and higher profitability.

@@

Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns. For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster fisheries were certified by the Marine Stewardship Council and lobster fisheries are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}. It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \citep{lester_2017}. Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access. This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change. Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be larger.

@@

While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat. Previous research focusing on these specific sites has shown that they serve as an insurance mechanism against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}. Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital as reported by \citet{fulton_2019}. Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}. These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-implementing the reserves \citep{ayer_2018}.

In terms of the fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs. Future research should aim at evaluating other Fish Refuges that have also been established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through top-down processes (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches.

Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}. We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside. Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness. For the particular case of the ones we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.

>>>>

---
title: "Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries"
rtitle: "Community-based TURF-reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,2}$, Stuart Fulton$^{3}$, Álvin Suarez$^{3}$, Arturo Hernández-Velasco$^{3}$, Jorge Torre$^{3}$, Fiorenza Micheli$^{4}$"
institution: |
  $^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA \newline
  $^{2}$Nicholas School of the Environment, Duke University, Beaufort, NC, USA \newline
  $^{3}$Comunidad y Biodiversidad A.C., Guaymas, Sonora, Mexico \newline
  $^{4}$Hopkins Marine Station and Center for Ocean Solutions, Stanford University, Pacific Grove, CA, USA
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "juancarlos@ucsb.edu"
abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last 12 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2014; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that while reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates, they continue to receive support from the fishing communities. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, some of these reserves might be too young for the effects to show. Second, the reserves are not large enough to protect mobile species, like lobster. Third, variable and extreme oceanographic conditions have impacted harvested populations. Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches. However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species."
keywords: "TURF-reserves, Causal Inference, Social-Ecological Systems, Marine Protected Areas, Marine Conservation, Small-Scale Fisheries"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

@@

TURFs are a fisheries management tool in which a well defined group of fishers (*e.g.* fishing cooperatives) have exclusive access to an explicitly delimited portion of the ocean. They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,costello_2010,mccay_2014}. On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools to aid in the recovery of marine stocks. These instruments can be combined by establishing a marine reserve within a TURF, thus making them TURF-reserves \citep{afflerbach_2014,gelcich_2015,lester_2017}.

Conservation science has shown how marine reserves lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits like climate change mitigation, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The benefits resulting from reserves established within TURFs (*i.e.* TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}. Although in theory these systems are successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success. Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{ostrom_2009,gelcich_2015}. It is therefore important to consider not only the indicators of interest, bu also the governance settings under which the reserves occur.

Recent norms in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries. In Mexico, a norm created in 2014 allows fishers to request legal recognition of community-based reserves as “Fish Refuges” (*Zona de Refugio Pesquero*; \citet{nom}). Since 2012, old and new marine reserves have gained legal recognition as Fish Refuges. Of these, 18 were originally implemented within TURFs. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

Here, we combine causal inference techniques and the Social-Ecological Systems (SES) framework to evaluate community-based TURF-reserves in three coastal communities in Mexico. The objective of this work is twofold. First, to provide a holistic evaluation of the effectiveness of community-based TURF-reserves in terms of the changes in biological and socioeconomic indicators and the governance settings under which these develop, which may inform similar processes in other countries. Second, to identify opportunities where improvement or adjustment might lead to increased effectiveness. We draw from lessons learned in these three case studies and provide management recommendations to maximize the effectiveness of community-based TURF-reserves in small-scale fisheries where this tool is used to manage and rebuild their coastal fisheries.

@@

In an effort to bridge this normative gap, Mexican Civil Society Organizations (CSOs) served as a link between fishers and government, and created a legal framework that solves this governance issue: Fish Refuges \citep{nom}. Fish Refuges can be implemented as temporal or partial reserves, which can protect one, some, or all resources within their boundaries. One of the ways in which fishing communities have taken advantage of this new tool is by implementing temporal marine reserves within their TURFs with a defined expiration date (often 5 years). When the expiration date is reached, fishers can chose to open the reserves to fishing or re-establish them. Our work focuses on Fish Refuges implemented as community-based TURF-reserves that occur in small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (*i.e.* TURFs shown as blue polygons in Fig \ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative. The profits from each TURF are shared amongst all fishers from the cooperative. Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (*i.e.* TURF-reserves). Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges following a series of studies outlined in the regulation \citep{nom}.

@@

We evaluate three community-based no-take TURF-reserves implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A). The first one was created by the *Buzos y Pescadores de la Baja California* fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the spiny lobster (*Panulirus interruptus*), but other resources like finfish, sea cucumber, read sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds. The objective of these reserves was "to protect commercially important invertebrate species"; mainly lobster and abalone. These reserves obtained legal recognition in 2018 \citep{dof_website_2018}.

The other two TURF-reserves are located in Maria Elena and Punta Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with Isla Nativdad, which hosts a well established fishing community, Maria Elena is a fishing camp --visited intermittently during the fishing season-- belonging to the *Cozumel* fishing cooperative; Punta Herrero is home to the *José María Azcorra* fishing cooperative, and similar to Isla Natividad hosts a local community. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off-season. Maria Elena and Punta Herrero established eight and four marine reserves in 2012 and 2013, respectively. These reserves have been legally recognized as Fishing Refuges since their original implementation \citep{dof_website_2012,dof_website_2013} and subsequent re-establishments \citep{dof_website_2017b}.

@@

The regulation governing the implementation of Fish Refuges states that these are fishery management tools intended to have biological or socioeconomic benefits \citep{nom}. For this reason, the main portion of our analyses focuses on a series of biological and socioeconomic indicators that may respond to reserve implementation. However, the effectiveness of conservation and fisheries management interventions also depends on the social and governance structures in place. We therefore incorporate a reduced version of the Social Ecological Systems framework \citep{ostrom_2009} and evaluate variables and indicators known to aid and hinder the effectiveness of management interventions in conservation and fisheries. The incorporation of the SES is not intended to relate different levels of governance with reserve effectiveness, but rather help provide context on the social-ecological system in which reserves develop. The following two sections describe our data collection methods and analyses.

@@

We use three main sources of information to evaluate these reserves across the ecological, socioeconomic, and governance dimensions. Ecological data come from the annual ecological monitoring of reserve and control sites. Reserve sites are areas where no fishing occurs. Control sites are areas that meet the following criteria: i) habitat characteristics are similar to the corresponding reserves, ii) presumably had a similar probability of being selected as reserves during the design phase, iii) are located within the TURF, where fishing occurs, and iv) Are not directly adjacent to the reserves. We focus our evaluation on sites where data are available for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{stewartoaten_1986,depalma_2018} and causally attribute the changes to the reserve \citep{francinifilho_2008,Villasenor-Derbez_2018}.

@@

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves. In this case our treated unit are the cooperatives (*i.e.* communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve. Cooperatives incorporated in this analysis belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2017,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}. A table with summary statistics for this data is provided in the supplementary materials (Table S3, Figure S5).

Data for the evaluation of the SES were collected at the community-level from official documents used in the design, creation, and designation of the marine reserves. These include the technical studies that the cooperatives submit when they request recognition of their reserves, as well as the official enactments  \citep{dof_website_2012,dof_website_2013,dof_website_2018}. We also complimented information based on the authors' experience and knowledge of the communities. We collected information on the Resource Systems, Resource Units, Actors, and Governance System (Table \ref{table:ses}). The next section further describe the variables and indicators chosen for each of these.

@@

\begin{equation}
I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Zone_i + \lambda_{t} Year_t\times Zone_i + \epsilon_{i,t}
\label{eqn:reg_bio}
\end{equation}

Where year-level fixed effects capturing a temporal trend are represented by $\gamma_t Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively). Therefore, we would expect this term to be positive if the indicator increases because of the reserve. Finally, $\epsilon_{i,t}$ represents the error term of the regression.

Socioeconomic indicators are evaluated with a similar approach. Due to data constrains, we only evaluate socioeconomic data for Isla Natividad (2000 - 2014) and Maria Elena (2006 - 2013). Neighboring communities are used as counterfactuals that allow us to control for unobserved time-invariants. Each focal community (*i.e.* Isla Natividad and Maria Elena) has three counterfactual communities.

\begin{equation}
I_{i,t} = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i +\epsilon_{i,t}
\label{eqn:soc_reg}
\end{equation}

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$). These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}. We fit one model per community and indicators (*e.g.* there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits. Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All analyses were performed in R `r paste0("version ", version$major, ".", version$minor)` and R Studio version 1.1.456 \citep{R_2018}. All data and code needed to reproduce our analyses are available in a GitHub repository at: [https://github.com/jcvdav/ReserveEffect](https://github.com/jcvdav/ReserveEffect).

@@

While many ecology studies have used BACI sampling designs and respective analyses (*e.g.* \citet{stewartoaten_1986}), few conservation studies have done so to evaluate the effect of an intervention (*e.g.* \citet{francinifilho_2008,lester_2009,moland_2013}) which has resulted in a call for more robust analyses in conservation science \citep{guidetti_2002,ferraro_2006}. Our approach to evaluate the temporal and spatial changes provides a more robust measure of reserve effectiveness, and captures previously described patterns. For example, the rapid increase observed for lobster densities in Isla Natividad on the sixth year (*i.e.* 2012; Fig. \ref{fig:indicators}A), occurs a year after the hypoxia events described by \citet{micheli_2012-EU}, which caused mass mortality of sedentary organisms such as abalone and sea urchins, but not lobster and finfish. The use of causal inference techniques may help us support evidence-based conservation.

Our analyses of socioeconomic indicators has two limitations. First, we only look at landings and revenues by landings for communities with and without TURF-reserves. There are a number of other possible indicators that could show a change due to the implementation of the reserve. Notably, one often cited in the literature is additional benefits, such as tourism \citep{viana_2017}. However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism. A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve. For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability.

@@

Even if reserves had appropriate sizes and were placed in optimal locations, there are other plausible explanations for the observed patterns. For instance, marine reserves are only likely to provide fisheries benefits if initial population sizes are low and the fishery is poorly managed \citep{hilborn_2004,hilborn_2006}. Both lobster fisheries were certified by the Marine Stewardship Council and are managed via species-specific minimum catch sizes, seasonal closures, protection of "berried" females, and escapement windows where traps are allowed \citep{dof_website_1993}. It is uncertain whether such a well-managed fishery will experience additional benefits from marine reserves; reserves implemented in TURFs where fishing pressure is already optimally managed will still show a trade-off between fisheries and conservation objectives \citep{lester_2017}. Furthermore, \citet{gelcich_2008} have shown that TURFs alone can have greater biomass and richness than areas operating under open access. This might reduce the difference between indicators from the TURF and reserve sites, making it difficult to detect such a small change. Further research should focus on evaluating sites in the reserve, TURF, and open access areas or similar Fish Refuges established without the presence of TURFs where the impact of the reserves might be greater.

@@

While the evaluated reserves have failed to provide fishery benefits up to now, there are a number of additional ecological, fisheries, and social benefits. Marine reserves provide protection to a wider range of species and vulnerable habitat. Previous research focusing on these specific sites has shown that they serve as an insurance mechanism against uncertainty and errors in fisheries management, as well as mild environmental shocks \citep{micheli_2012-EU,deleo_2015,roberts_2017-J9,aalto}. Self-regulation of fishing effort can serve as a way to compensate for future declines associated to environmental variation \citep{finkbeiner_2018}. Furthermore, embarking in a marine conservation project can bring the community together, which promotes social cohesion and builds social capital \citep{fulton_2019}. Showing commitment to marine conservation and sustainable fishing practices has allowed fishers to have greater bargaining power and leverage over fisheries management \citep{prezramrez_2012}. These additional benefits might explain why communities show a positive perception about their performance and continue to support their presence by re-implementing the reserves \citep{ayer_2018}.

Community-based TURF-reserves in small-scale fisheries may be helpful conservation and fishery management tools when appropriately implemented \citep{gelcich_2015}. We must promote bottom-up design and implementation processes like the ones in the evaluated reserves, but without setting design principles aside. Having full community support surely represents an advantage, but it is important that community-based TURF-reserves meet essential design principles such as size and placement so as to maximize their effectiveness. Furthermore, consvation and advocacy groups should consider the opportunity costs of such interventions (*sensu* \citet{smith_2010}) and evaluate the potential of other approaches that may yield similar benefits.

In terms of the fisheries regulation in Mexico, our work only evaluates Fish Refuges established within TURFs. Future research should aim at evaluating other Fish Refuges established as bottom-up processes but without the presence of TURFs (*e.g.* \citet{dof_websiteC_2012}), others established through top-down processes (*i.e.* \citet{dof_websiteU_2018}), as well as the relationship between governance and effectiveness across this gradient of approaches. For the particular case of the reserves that we evaluate, the possibility of expanding reserves or merging existing polygons into larger areas should be evaluated and proposed to the communities.
